---
title: Instructions to create `json` files for `GeoplotR`
author: Pieter Vermeesch
output:
  pdf_document:
    includes:
      keep_tex: yes
    number_sections: yes
---

```{r settings, include = FALSE}
# rmarkdown::render("~/Documents/Programming/R/GeoplotR/helpers/instructions.Rmd")
```

# Getting started

1. Install `R` and `RStudio` from:
   
   [`https://rstudio.com/products/rstudio/download/#download`](https://rstudio.com/products/rstudio/download/#download)

1. Start `RStudio` and install the `remotes` package by typing the
following code at the console (i.e. the '`>`'-symbol):

   ```{r eval=FALSE}
   install.packages('remotes')
   ```

1. Now install the latest beta version of the `GeoplotR` package by
typing the following commands at the console:

   ```{r eval=FALSE}
   remotes::install_github('https://github.com/pvermees/GeoplotR/')
   ```

1. Download `test.R`, `Template.json`, `Cr_Y.json`, `QAP.json` from
[`github:pvermees/GeoplotR/helpers/`](https://github.com/pvermees/GeoplotR/blob/main/helpers/)
and save them in a folder on your computer.

# Using the ```Excel` to `json`

1. You can open the `json` files in `RStudio`. They contain the
following information:

   (a) `lines`: the coordinates of anchor points on a decision boundary

   (b) `line_type`: `solid`, `dashed` or `dotted` (one for each item in `lines`)

   (c) `polygons`: the coordinates of the corners of polygons whose edges
   overlap with the entries in `lines`

   (d) `label_coords`: the coordinates of the labels that are to be
   placed inside each of the polygons.

   (e) `labels`: the actual labels (one for each item in `label_coords`)

   (f) `angle`: the rotation angle of the labels (one for each item
   `label_coords`)

1. Open `test.R` in `RStudio` and click on the `source` button. This
should create a PDF document on your computer, in the same folder as
the `test.R` file. Inspect the contents of the PDF document to
understand what `test.R` has done.

1. Have a look at `Template.json` in `RStudio`:

   (a) make any change to any number in this file

   (b) save the file

   (c) go back to the `test.R` tab

   (d) click on `source` again.

   Do you notice any change?

1. `test.R` contains computer code in the `R` language. It uses two
functions of the `GeoplotR` package named `xytest` and `xyztest`,
respectively.  You can read more about these functions, and about
their arguments, in the built-in documentation, by typing the
following code at the console:

   ```{r eval=FALSE}
   ?xyplot
   ?xyzplot
   ```

# `Excel` to `json`

1. Visit
[`https://minerva.union.edu/hollochk/c_petrology/discrim.html`](https://minerva.union.edu/hollochk/c_petrology/discrim.html)
and click on any blue link in it. This should download an `Excel` file
on your computer containing the anchor points of various binary and
ternary diagrams.

1. Your job is to use the coordinates in these `Excel` files to
produce lines, polygons and labels that reproduce the various plots on
the website. I have already finished implementing several diagrams,
whose `json` files you can inspect at
[github:pvermees/GeoplotR/inst/](https://github.com/pvermees/GeoplotR/tree/main/inst). I
would like you to make some more, notably:

   * Meschede, 1986

   * Pearce and Cann, 1973 (except the Ti-Zr-Y diagram, which has
     already been implemented)

   * Wood, 1980

   * Ewart, 1982

   * Winchester and Floyd, 1977

   * Cabanis and Lecolle, 1989

   * Mullen, 1983

   * Shervais, 1982

   * Pearce and Norry, 1979

1. The best way to build a `json` file is to open `Template.json`,
save it under a different name and modify its content. Edit `test.R`
to test your `json` file. Test frequently to catch bugs! So: edit,
test, edit, test, ... until the diagram looks like the one in the
`Excel` file.